---
title: "R codes for the paper *Covariates impacts in spatial autoregressive models for compositional data*"
date: "Last update: `r Sys.Date()`"
author: C. Thomas-Agnan, T. Laurent and A. Ruiz-Gazen
header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
output:
  pdf_document:   
  number_sections: yes
---
  
  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(fig.align = "center",
               cache = F,
               fig.width = 7,
               fig.height = 6,
               cache.lazy = FALSE
)
options(kableExtra.latex.load_packages = FALSE)
```

This document presents the **R** codes used to obtain the computational results included in the paper "Covariates impacts in spatial autoregressive models for compositional data". 
To cite this work, please use:
  
C. Thomas-Agnan, T. Laurent and A. Ruiz-Gazen (2023). Covariates impacts in spatial autoregressive models for compositional data, *WP*.

Packages needed:
  
```{r, message = F, warning = F}
library(cartography) # mapping
library(compositions) # compositions data
library(gridExtra) # several frame in ggplot
library(Matrix)    # sparse matrix
library(scatterpie) # plot several pie 
library(sp) # spatial package sp
library(spdep) # spatial econometrics
library(sf) # spatial package sf
library(tidyverse) # tidyverse universe 
```



# Section "Impacts decomposition and summaries" in the article

## Create the data

We consider the $n=13$ departments of the Occitanie region in France whose geometries have been simplified. We simulate the unemployment rate using an uniform law with parameters $[0,0.3]$.

```{r}
n_occ <- 13
id_region_occ <- c("LT", "LZ", "TG", "TA", "AV", "GA", 
                   "GE", "HG", "AU", "HE", "HP", "AR", "PO")
set.seed(12345)
occ_df <- data.frame(id = id_region_occ,
                     unemp = runif(n_occ, 0, 0.3),
                     row.names = "id")

sp_occ <- SpatialPointsDataFrame(cbind(c(0, 0, 0, 0, 1, 1, 1, 2, 2, 2, 2, 3, 3),
                                       c(0, 1, 2, 3, 0, 1, 2, 0, 1, 2, 3, 1, 2)),
                                 occ_df)

create_grid <- function(sp_centroid) {
  # initialization
  # define the grid
  coords <- coordinates(sp_centroid)
  id <- row.names(sp_centroid)
  nx <- length(unique(coords[, 1])) 
  ny <- length(unique(coords[, 2]))  
  gt <- GridTopology(c(min(coords[, 1]), min(coords[, 2])), c(1, 1), c(nx, ny))
  grd <- SpatialGrid(gt)
  spix <- as(grd, "SpatialPixels")
  spol <- as(spix, "SpatialPolygons")
  o <- over(spol, sp_centroid)[, 1]
  simu_spdf <- spol[!is.na(o), ]
  row.names(simu_spdf) <- id
  simu_spdf <- SpatialPolygonsDataFrame(simu_spdf, sp_centroid@data) 
  return(simu_spdf)
}

spdf_occ <- create_grid(sp_centroid = sp_occ)
sf_occ <- as(spdf_occ, "sf")
```


## Plotting the data

We represent the map of the "simulated" unemployment rate (not presented in the article):
  
```{r, echo=F, fig.height = 5, fig.width = 5}
# pdf(file = "./figures/explanatory.pdf", width = 14, height = 8)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
# plot municipalities (only the backgroung color is plotted)
plot(st_geometry(sf_occ), col = NA, border = NA, bg = "#aadaff")
# plot population density
choroLayer(
  x = sf_occ, 
  var = "unemp",
  method = "kmeans",
  col = carto.pal(pal1 = "sand.pal", n1 = 5),
  border = "white", 
  lwd = 0.5,
  legend.values.rnd = 3, 
  legend.pos = "topright", 
  legend.title.txt = "Unemployment rate",
  add = TRUE
) 
# layout
layoutLayer(title = "Unemployment rate in Occitanie", 
            sources = "Source: Insee 2020",
            frame = FALSE, north = FALSE, tabtitle = TRUE, 
            theme= "sand.pal", scale = F) 
text(coordinates(spdf_occ)[, 1], coordinates(spdf_occ)[, 2], row.names(occ_df))
```


## Initialization

Before simulating the spatial CoDa model:
  
* we define a spatial weight matrix based on the rook-contiguity (the plot was not presented in the article)

```{r, fig.height = 5, fig.width = 5}
my_nb <- poly2nb(spdf_occ, queen = F)
W <- listw2mat(nb2listw(my_nb))
# pdf(file = "./figures/w.pdf", width = 8, height = 8)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot(spdf_occ)
plot(my_nb, coordinates(spdf_occ), col = "grey", add = T)
# dev.off()
```

## Simulation

We simulate a $n\times 2$ matrix $\mathbf{Y}^*$ of the form (the choice of the coefficients have been inspired by the results of the spatial multivariate regression model found in Nguyen *et al.*, 2020)

$$
  \mathbf{Y}^*= \pmb{\mu}^{*} + unemp\times \pmb{\beta}^{*}+ \mathbf{W}  \mathbf{Y}^* \mathbf{R}^* + \pmb{\epsilon}^*.
$$
with $\pmb{\mu}^{*}=(0.0696, -0.312)$, $\pmb{\beta}^{*}=(-0.88,1.82)$, $\mathbf{R}^*=\begin{bmatrix}
0.65  &  0  \\
0.18  &  0.63 
\end{bmatrix}$ and $\mathbf{\Sigma}^*=\begin{bmatrix}
0.09  &  0.02  \\
0.02  &  0.09
\end{bmatrix}.$ 
  
We source the function that allows to simulate a spatial multivariate process (the function is described in http://author/code/spatial_coda/):
  
```{r, message = F, eval = F}
source(url("http://author/code/spatial_coda/R/simu_spatial_multi_y.R"))
```

```{r, message = F, echo = F}
source(url("http://www.thibault.laurent.free.fr/code/spatial_coda/R/simu_spatial_multi_y.R"))
```



```{r, message = F}
my_x <- cbind(1, as.matrix(occ_df[, "unemp"]))
set.seed(12345)
y_star <- simu_spatial_multi_y(X = my_x, 
                               beta_true = matrix(c(0.0696, -0.88, 
                                                    -0.312, 1.82), 
                                                  byrow = F, ncol = 2), 
                               Sigma = matrix(c(0.09, 0.02, 0.02, 0.09), 2, 2), 
                               RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), 
                               W = W) 
```

We transform the $Y^*$ matrix back to the simplex for representing the three components:
  
```{r}
Y <- ilrInv(y_star)
sf_occ[, c("y1", "y2", "y3")] <- Y
```

We represent the map of the different components of $Y$. The function *my_rgb_coda()* allows to choose appropriate colors:
  
  
```{r, echo = F}
source("R/my_rgb_coda.R")
```

```{r, eval = F}
source(url("http://www.thibault.laurent.free.fr/code/sp_coda_impacts/R/my_rgb_coda.R"))
```

```{r, eval = F}
source(url("http://author/code/sp_coda_impacts/R/my_rgb_coda.R"))
```

We present here some other alternatives for mapping compositional data. It can be particularly interesting if $D>3$. First alternative is the waffle layer, second is the ternary diagram. 
  
```{r, fig.width = 10, fig.height = 6}
sf_occ$y1_100 <- round(sf_occ$y1 * 100)
sf_occ$y2_100 <- round(sf_occ$y2 * 100)
sf_occ$y3_100 <- round(sf_occ$y3 * 100)

sf_occ$y2_100[10] <- sf_occ$y2_100[10] - 1
sf_occ$y1_100[12] <- sf_occ$y1_100[12] + 1

R <- c(0, 0)
G <- c(1, 0)
B <- c(0.5, sqrt(3) / 2)
# create a grid of shares for Y
my_don_x_new <- Y[, 1] * R[1] + Y[, 2] * G[1] + Y[, 3] * B[1] 
my_don_y_new  <- Y[, 1] * R[2] + Y[, 2] * G[2] + Y[, 3] * B[2] 
```

To get Figure 1 in the article : 

```{r, fig.width = 10.5, fig.height = 3.5}
# pdf(file = "./figures/ternary_to_ilr.pdf", width = 10.5, height = 3.5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), mfrow = c(1, 3))
plot(st_geometry(sf_occ), col = "#f2efe9", border = "#b38e43", lwd = 0.5)
waffleLayer(
  x = sf_occ,
  var = c("y1_100", "y2_100", "y3_100"),
  cellvalue = 1,
  cellsize = 0.04,
  cellrnd = "ceiling",
  celltxt = "",
  labels = c("Left", "Right", "XR"),
  ncols = 10,
  col = my_rgb_coda(as(rbind(c(1, 0, 0), c(0, 1, 0), c(0, 0, 1)), "matrix"), 
                    max_intensity = 1.5),
  border = "#f2efe9",
  legend.pos = "topright",
  legend.title.cex = 1,
  legend.title.txt = "Simulated \n vote shares",
  legend.values.cex = 1,
  add = TRUE
)
coords_dep <- st_coordinates(st_centroid(sf_occ))
text(coords_dep[, 1], coords_dep[, 2]- 0.35, 
     paste0(c("Lot", "Lozère", "Tarn-et-\nGaronne", "Tarn", "Aveyron", 
       "Gard", "Gers", "Haute-\nGaronne", "Aude", "Hérault", 
       "Hautes-\nPyrnées", "Ariège", "Pyrénées-\nOrientales"), 
       paste0(" (", 1:13, ")")), cex = 0.9
)

plot(rbind(R, G, B), 
     xaxt = "n", 
     yaxt = "n", 
     frame = F, 
     pch = 16, 
     type = "n",
     xlab = "", 
     ylab = "", 
     asp = 1, 
     ylim = c(-0.1, sqrt(3)/2)
)

lines(c(0, 1), c(0, 0), col = "black")
lines(c(0, 0.5), c(0, sqrt(3)/2), col = "black")
lines(c(1, 0.5), c(0, sqrt(3)/2), col = "black")

base_trian <- (G[1] - R[1]) 
hauteur <- B[2] - R[2]

# XR lines
for (k in 1:4) {
  lines(c(R[1] + base_trian * k/10, G[1] - base_trian * k/10), 
        c(R[2] + k/5 * hauteur, R[2] + k/5 * hauteur), lty = 2, lwd = 0.8)
  text(G[1] - base_trian * k/10, R[2] + k/5 * hauteur, 2 * k/10, pos = 4, cex = 1)
}
text(G[1] - base_trian * 1/2, R[2] + hauteur, "1 = XR", pos = 4, cex = 1)

# Left lines
for (k in 1:4) {
  lines(c(R[1] + k/5, base_trian / 2 + k * 1/10), 
        c(R[2], B[2] - k/5 * hauteur), lty = 2, lwd = 0.8)
  text(R[1] + k/5 * base_trian, R[2] - 0.02, 2 * k/10, pos = 1, cex = 1, srt = 70)
}
text(R[1] + base_trian, R[2], "1 = Right", pos = 1, cex = 1, srt = 70)
  
# Right lines 
for (k in 1:4) {
  lines(c(R[1] + base_trian * k * 1 / 10, R[1] + base_trian * k/5), 
        c(R[2] + k/5 * hauteur, R[2]), lty = 2, lwd = 0.8)
  text(R[1] + base_trian * k * 1 / 10, R[2] + k/5 * hauteur, 1 - 2 * k/10, 
       pos = 2, cex = 1, srt = -50)
}
text(R[1]-0.04,  R[2] + 0.02, "1 = Left", 
       pos = 4, cex = 1, srt = -50)
  
points(my_don_x_new, my_don_y_new, pch = 16, 
       col = my_rgb_coda(as(Y, "matrix"), max_intensity = 1.5), 
       cex = 1)
#text(c(0.04, 0.95, 1/2), c(0-0.06, -0.06, sqrt(3)/2 - 0.005), 
#     c("Left", "Right", "XR"), pos = c(2, 4, 3), 
#     cex = 0.7)
# map
plot(st_geometry(sf_occ), col = my_rgb_coda(as(Y, "matrix"), max_intensity = 1.5))
# dev.off()
```



Second alternative is based on the pie (figure not presented in the article): 
  
```{r, warning = F, message = F, fig.height = 5, fig.width = 5}
sf_occ[ , c("y1", "y2", "y3")] <- as(Y, "matrix")
sf_occ[ , c("long", "lat")] <- st_coordinates(st_centroid(sf_occ))
sf_occ$region <- row.names(sf_occ)
ggplot()  + 
  geom_scatterpie(aes(x = long, y = lat, group = region), data = st_drop_geometry(sf_occ),
                  cols = c("y1", "y2", "y3"), pie_scale = 4) + 
  coord_equal() + 
  geom_sf(data = sf_occ, fill = NA)
```

Summary statistics (not presented in the article)

```{r}
kableExtra::kbl(round(t(sapply(
  st_drop_geometry(sf_occ)[ , c("unemp", "y1", "y2", "y3")], 
  function(x) c(min = min(x), max = max(x), 
                mean = mean(x), median = median(x),  
                sd = sd(x)))), 3))
```

## Computing the semi-elasticities

The function *sp_semi_elast_Ycompo()* computes the semi-elasticities individual by individual in the spatial-compositional regression model. It takes as input arguments:
  
* **X**, the covariates matrix of size $n\times p$ (with the intercept if included in the model)
* **b_star** the matrix of the estimated parameter $\hat\beta^*$ of size $p\times (D-1)$, in the spatial model in the ILR space, 
* **index_b**, an integer, the column index of the variable that will be considered to compute the semi-elasticities 
* **W**, the spatial weight matrix of the $n$ observations
* **RHO** the matrix of the estimated spatial autocorrelation parameter
* **type**, a character with **"exact"** or **"appro"** that indicated the method to be used
* **print_summary**, a logical, if TRUE, print the summary

```{r, echo = F}
source("R/sp_semi_elast_Ycompo.R")
```

```{r, eval = F}
source(url("http://author/code/sp_coda_impacts/R/sp_semi_elast_Ycompo.R"))
```

```{r, eval = F, echo = F}
source(url("http://www.thibault.laurent.free.fr/code/sp_coda_impacts/R/sp_semi_elast_Ycompo.R"))
```

**Application:** we use function *sp_semi_elast_Ycompo()* to compute the semi-elasticities on our simulated data:

```{r}
my_beta <- matrix(c(0.0696, -0.88, -0.312, 1.82),
                  byrow = F, ncol = 2)
elast_sp <- sp_semi_elast_Ycompo(my_x, my_beta, index_b = 2, 
                                 method = "appro", W = W, 
                                 RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2))
```


## Representing the semi-elasticities

We can map the matrices of semi-elasticities

```{r}
rwb <- colorRampPalette(colors = c("blue", "white", "red"))
bks <- seq(-max(abs(range(as.vector(elast_sp)))), 
           max(abs(range(as.vector(elast_sp)))), length.out = 21)
```

For instance we map the impacts due to observation 5 on all the network:

```{r, warning = F, fig.width=12,fig.height=4.5}
k <- 5
#pdf("figures/impacts_s5.pdf", width = 10.5, height = 3.5)
op <- par(mfrow = c(1, 3), oma = c(0, 0.5, 1, 0), mar = c(0, 0.5, 1, 0), 
          mgp = c(1, 1, 1))
plot(st_geometry(sf_occ), col = rwb(20)[findInterval(elast_sp[, k, 1], 
     bks, all.inside = T)], main = "s.e. Left", 
     ylab = paste0("Impacts due to change of x at s", k))

plot(st_geometry(sf_occ[5, ]), add = T, lwd = 3, lty = 2)

text(st_coordinates(st_centroid(sf_occ))[, 1], st_coordinates(st_centroid(sf_occ))[, 2], 
      paste0(paste("se(", 1:13,",", 5, ")", sep = ""), "\n =", 
             round(elast_sp[, k, 1], 2)))
text(st_coordinates(st_centroid(sf_occ))[, 1] - 0.08, 
     st_coordinates(st_centroid(sf_occ))[, 2]+ 0.04, "1", 
     cex = 0.7, pos = 3) 
plot(st_geometry(sf_occ), col = rwb(20)[findInterval(elast_sp[, k, 2],
                            bks, all.inside = T)],
     main = "s.e. Right")
text(st_coordinates(st_centroid(sf_occ))[, 1], st_coordinates(st_centroid(sf_occ))[, 2], 
      paste0(paste("se(", 1:13,",", 5, ")", sep = ""), "\n =", 
             round(elast_sp[, k, 2], 2))) 
text(st_coordinates(st_centroid(sf_occ))[, 1] - 0.1, 
     st_coordinates(st_centroid(sf_occ))[, 2]+ 0.04, "2", 
     cex = 0.7, pos = 3) 
plot(st_geometry(sf_occ[5, ]), add = T, lwd = 3, lty = 2)

plot(st_geometry(sf_occ), col = rwb(20)[findInterval(elast_sp[, k, 3], 
                                         bks, all.inside = T)],
     main = "s.e. XR")
text(st_coordinates(st_centroid(sf_occ))[, 1], st_coordinates(st_centroid(sf_occ))[, 2], 
      paste0(paste("se(", 1:13,",", 5, ")", sep = ""), "\n =", 
             round(elast_sp[, k, 3], 2)))
text(st_coordinates(st_centroid(sf_occ))[, 1] - 0.08, 
     st_coordinates(st_centroid(sf_occ))[, 2]+ 0.04, "3", 
     cex = 0.7, pos = 3) 
plot(st_geometry(sf_occ[5, ]), add = T, lwd = 3, lty = 2)

par(op)
#dev.off()
```

With the follwing color codes:
```{r, fig.width = 12, fig.height = 1.5}
#pdf("figures/se_colors.pdf", width = 12, height = 1.5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot(c(0, 1), c(0, 0), frame.plot = F, xaxt = "n", 
     yaxt = "n", ylab = "", pch = "|",
     xlab = "", xlim = c(-2.05, 2.05))
title("Semi-elasticities", line = -2)
for(k in 1:(length(bks) - 1)) {
  polygon(c(bks[k], bks[k+1], bks[k+1], bks[k], bks[k]),
          c(-0.28, -0.28, 0.08, 0.08, -0.28),
          col = rev(rwb(20))[k])
  text(bks[k], 0, round(bks[k], 2), pos = 3)
}
  text(bks[k+1], 0, round(bks[k+1], 2), pos = 3)
#dev.off()
```



We can map the impacts to each observation (not presented in the article) :

```{r, fig.width=20, fig.height = 60, warning = F, message = F}
op <- par(mfrow = c(13, 3), oma = c(0, 2, 2, 0), mar = c(0, 2, 2, 0), 
          mgp = c(1, 1, 1))
for(k in 1:13) {
 plot(st_geometry(sf_occ), col = rwb(20)[findInterval(elast_sp[, k, 1], 
                                            bks, all.inside = T)],
     main = "s.e. Left", ylab = paste0("Impacts due to change of x at s", k))
 text(st_coordinates(st_centroid(sf_occ))[, 1], st_coordinates(st_centroid(sf_occ))[, 2], 
      paste0(paste("se(", 1:13, ",", k, ")", sep = ""), "\n =", 
             round(elast_sp[, k, 1], 2)))   
 plot(st_geometry(sf_occ), col = rwb(20)[findInterval(elast_sp[k, , 2], 
                                          bks, all.inside = T)],
     main = "s.e. Right")
 text(st_coordinates(st_centroid(sf_occ))[, 1], st_coordinates(st_centroid(sf_occ))[, 2], 
      paste0(paste("se(", 1:13, ",", k, ")", sep = ""), "\n =", 
             round(elast_sp[, k, 2], 2)))  
 plot(st_geometry(sf_occ), col = rwb(20)[findInterval(elast_sp[, k, 3], 
                                                bks, all.inside = T)],
     main = "s.e. XR")
 text(st_coordinates(st_centroid(sf_occ))[, 1], st_coordinates(st_centroid(sf_occ))[, 2], 
      paste0(paste("se(", 1:13, ",", k, ")", sep = ""), "\n =", 
             round(elast_sp[, k, 3], 2))) 
}
par(op)
```


We now plot the boxplot of the elasticities with respect to their group: direct, indirect

```{r}
D <- 3
res_boxplot <- data.frame(
  obs = rep(rep(sf_occ$region, each = n_occ), times = D),
  elas = c(as.vector(elast_sp[ , , 1]), 
           as.vector(elast_sp[ , , 2]), 
           as.vector(elast_sp[ , , 3])),
  type = ifelse(as.vector(diag(n_occ)) == 1, "direct", "indirect"),
  comp = rep(c("Left", "Right", "XR"), each = n_occ ^ 2))
```

```{r, fig.width = 6, fig.height = 4}
ggplot(data = res_boxplot) +
  aes(y = elas, fill = type, x = comp) +
  geom_boxplot() +
  xlab("Component") + 
  ylab("semi-elasticities")
#ggsave("figures/boxplot_semi.pdf", width = 6, height = 4)
```


We plot the density plot of the semi-elasticities:

```{r, fig.width = 6, fig.height = 4}
ggplot(data = res_boxplot) +
  aes(x = elas, fill = type) +
  geom_density() +
  facet_wrap(~comp, ncol = 1) +
  xlab("semi-elasticities") + 
  ylab("")
# ggsave("figures/density_semi.pdf", width = 5, height = 4)
```

## Groupping the semi-elasticities

The function *map_elas_ternary()* takes as argument :

* **df_coords** the coordinates of the spatial observations
* **X**, the covariates matrix of size $n\times p$ (with the intercept if included in the model)
* **b_star** the matrix of the estimated parameter $\hat\beta^*$ in the spatial model in the ILR coordinates, of size $p\times (D-1)$
* **index_b**, an integer, the column index of the variable that will be considered to compute the semi-elasticities 
* **W**, the spatial weight matrix of the $n$ observations
* **RHO** the matrix of the estimated spatial autocorrelation parameter
* **delta**, the value of the increment to compute the $\hat Y(x+\delta)$
* **type**, a integer equals to 1 or 2. The first case focuses on the impacts aggregated by spatial unit whereas the second one illustrates the impacts due a spatial unit on all others 
* **scale_ternary**, a logical, if TRUE, plot the scale on the ternary diagram
* **add**, a logical, if TRUE, do not open a new graphic device
* **range**, corresponds to $x_{max}-x_{min}$ where $x_{min}$ is the minimum of the $x$-geographical coordinate and $x_{max}$ is the maximum of the $x$-geographical coordinate
* **cex.pts**, the size of the points 
* **size**, a scalar that fits the size of the ternary diagram
* **length**, the size of the arrow 
* **cex.legend**, the size of the police 
* **legend.ternary**, a vector of size 3 of characters that give the legend for the ternary diagram. 
* **ternary_unit**, a logical, if TRUE prints the scaling of the ternary diagram

```{r, echo = F}
source("R/map_elas_ternary.R")
```

```{r, eval = F}
source(url("http://author/code/sp_coda_impacts/R/map_elas_ternary.R"))
```

```{r, eval = F, echo = F}
source(url("http://www.thibault.laurent.free.fr/code/sp_coda_impacts/R/map_elas_ternary.R"))
```

```{r, warning = F}
coords_s <- st_coordinates(st_centroid(sf_occ))
rownames(coords_s) <- sf_occ$region
```


# Alternative 1

We propose to make a zoom on the direct $DE(i)$ and indirect $IE(i)$ observed at each location $i$. 

In red, we represent the total impact, in red the direct impact and in green the indirect impact. 

```{r, warning = F,fig.width = 5, fig.height = 5}
#pdf("figures/local_v2.pdf", width = 6, height = 6)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot(st_geometry(sf_occ), col = rgb(0.9, 0.9, 0.9), border = "white")
coords_s_ts <- coords_s 
coords_s_ts[, 2] <- coords_s_ts[, 2] - 0.05
map_elas_ternary(coords_s_ts, 
                  my_x, my_beta, index_b = 2, W = W,
                  RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), delta = 0.3,
                  range = 3, size = 2.7, lwd = 1.5,  length = 0.01, cex.legend = 0.65,
                              legend.ternary = c("L", "R", "XR"))
plot(st_geometry(sf_occ[5, ]), add = T, lwd = 1, lty = 2)

legend("topright", legend = c("direct", "indirect"), pch = c(NA, NA),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, box.lwd = -1) #normal legend. Do not plot line
par(font = 5) #change font to get arrows
legend("topright", legend = c("        ", "          "), pch = c(174, 174),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, bty = "n", box.lwd = -1) 
#dev.off()
```

```{r, fig.width = 5, fig.height = 5}
#pdf("figures/zoom_aveyron_2.pdf", width = 5, height = 5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot(st_geometry(sf_occ[5, ]), border = "white", ylim = c(1.7, 2.5))
map_elas_ternary(coords_s, add = T, row_to_plot = 5,
                  X = my_x, b_star = my_beta, index_b = 2, W = W,
                 type = 1, scale_ternary = TRUE,
                  RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), 
                 delta = 0.3,
                  range = 3, cex.pts = 0.2, size = 3.2, 
                 length = 0.03, 
                 cex.legend = 0.7, lwd = 2,
                 legend.ternary = c("L", "R", "XR"))
legend("topright", legend = c("direct", "indirect"), pch = c(NA, NA),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, box.lwd = -1) #normal legend. Do not plot line
par(font = 5) #change font to get arrows
legend("topright", legend = c("        ", "          "), pch = c(174, 174),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, bty = "n", box.lwd = -1) 
#dev.off()
```



### Alternative 2

We aim to plot the impacts caused by changing $X$ at one particular observation. 

```{r, warning = F,fig.width = 5, fig.height = 5}
#pdf("figures/local_v1.pdf", width = 6, height = 6)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
coords_s_ts <- coords_s 
coords_s_ts[, 2] <- coords_s_ts[, 2] - 0.05
plot(st_geometry(sf_occ), col = rgb(0.9, 0.9, 0.9), border = "white")
map_elas_ternary(df_coords = coords_s_ts, 
                  X = my_x, b_star = my_beta, index_b = 2, W = W,
                 type = 2,
                  RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), delta = 0.3,
                  range = 3, cex.pts = 0.2, size = 2.7, lwd = 1.5,  length = 0.01,
                 cex.legend = 0.65, legend.ternary = c("L", "R", "XR"))
plot(st_geometry(sf_occ[5, ]), add = T, lwd = 1, lty = 2)
legend("topright", legend = c("direct", "indirect"), pch = c(NA, NA),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, box.lwd = -1) #normal legend. Do not plot line
par(font = 5) #change font to get arrows
legend("topright", legend = c("        ", "          "), pch = c(174, 174),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, bty = "n", box.lwd = -1) 
#dev.off()
```

We zoom on the Aveyron departement:

```{r, fig.width = 5, fig.height = 5}
#pdf("figures/zoom_aveyron_1.pdf", width = 5, height = 5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot(st_geometry(sf_occ[5, ]), border = "white", ylim = c(1.7, 2.5))
map_elas_ternary(coords_s, add = T, row_to_plot = 5,
                  X = my_x, b_star = my_beta, index_b = 2, W = W,
                 type = 2, scale_ternary = TRUE,
                  RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), delta = 0.3,
                  range = 3, cex.pts = 0.2, size = 3.2, 
                 length = 0.03, 
                 cex.legend = 0.7, lwd = 2,
                 legend.ternary = c("L", "R", "XR"))
legend("topright", legend = c("direct", "indirect"), pch = c(NA, NA),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, box.lwd = -1) #normal legend. Do not plot line
par(font = 5) #change font to get arrows
legend("topright", legend = c("        ", "          "), pch = c(174, 174),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, bty = "n", box.lwd = -1) 
#dev.off()
```



## Aggregating the semi-elasticities into local direct and indirect impacts

We summarize the results into direct, indirect and total effects:

```{r}
spatial_direct <- apply(elast_sp, 3, function(x) diag(x))
spatial_total <- apply(elast_sp, c(1, 3), sum)
spatial_indirect <- spatial_total - spatial_direct
colnames(spatial_direct) <- colnames(spatial_total) <- 
  colnames(spatial_indirect) <- c("Left", "Right", "XR")
```


### Barplot of the semi-elasticities

We can map the three impacts $DE(i),IE(i),TE(i)$ at each location $i$ on the same map. We have implemented function *map_elas()* that allows to print the local effects. The imput arguments are : 

* **spatial_direct**, **spatial_indirect**, **spatial_total**, the matrices of size $n\times D$ of the direct, indirect and total impacts
* **coords_s**, the matrices of size $n\times 2$ of the spatial coordinates. **row.names** must be given to print the legend
* **q4**, the vector of colors for each component $d \in D$. The ordering is the same than the columns of **spatial_direct**
* **add**, a logical indicating whether a new graphic window may be opened.
* **max_bar** the value of the highest bar of the bar plots
* **max_width** the value of the width of the bars 
* **x.legend**, a character among "none", "bottomright", "bottom", "bottomleft", 
            "left", "topleft", "top", "topright", "right", "center" that gives the location of the legend
* **label_s**, a logical for printing the names of the entities 
* **plot_scale**, a logical, if TRUE prints the scaling

```{r, echo = F}
source("R/map_elas_bar.R")
```

```{r, eval = F, echo = F}
source(url("http://www.thibault.laurent.free.fr/code/sp_coda_impacts/R/map_elas_bar.R"))
```

```{r, eval = F}
source(url("http://author/code/sp_coda_impacts/R/map_elas_bar.R"))
```


```{r, warning = F}
coords_s <- st_coordinates(st_centroid(sf_occ))
rownames(coords_s) <- sf_occ$region
```

```{r, fig.width = 5, fig.height = 5}
#pdf("figures/bar_local.pdf", width = 5, height = 5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot(st_geometry(sf_occ), 
     col = rgb(0.96, 0.96, 0.96), border = "white")
map_elas_bar(spatial_direct, spatial_indirect, spatial_total, coords_s, add = T,
          q4 = my_rgb_coda(as(rbind(c(1, 0, 0), c(0, 1, 0), c(0, 0, 1)), "matrix"), 
                    max_intensity = 1.5),
         width_bar = 4, max_bar = 4, cex.legend = 0.65, x.legend = "topright",
         label_s = F, round.scale = 0)
#dev.off()
```


### Ternary diagram of the local impacts 

We aggregate the indirect impacts and plot them in the ternary diagram 

```{r, warning = F, fig.width = 5, fig.height = 5}
#pdf("figures/local_v3.pdf", width = 5, height = 5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot(st_geometry(sf_occ), col = rgb(0.9, 0.9, 0.9), border = "white")
map_elas_ternary(coords_s_ts, 
                  my_x, my_beta, index_b = 2, W = W, all_impacts = F,
                  RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), delta = 0.3,
                  range = 3, size = 2.7, lwd = 1.5,  length = 0.01, cex.legend = 0.5,
                              legend.ternary = c("L", "R", "XR"))

#legend("topright", legend = c("direct", "indirect", "total"), lty = 1,
#        col = c("#E16A86", "#009ADE", "#50A315"), cex = 0.9, box.lwd = -1)

legend("topright", legend = c("direct", "indirect", "total"), pch = c(NA, NA, NA),
       lwd = 3, col = c("#E16A86", "#009ADE", "#50A315"),
       cex = 0.9, lty = NA, box.lwd = -1) #normal legend. Do not plot line
par(font = 5) #change font to get arrows
legend("topright", legend = c("        ", "          ", "       "), pch = c(174, 174, 174),
       lwd = 3, col = c("#E16A86", "#009ADE", "#50A315"),
       cex = 0.9, lty = NA, bty = "n", box.lwd = -1) 

#dev.off()
```



# Case study 

## Preparation of the data 

We first import the data: 

```{r, message = F, warning = F, echo = F}
load(url("http://www.thibault.laurent.free.fr/code/spatial_coda/R/data_cantons.RData"))
source(url("http://www.thibault.laurent.free.fr/code/spatial_coda/R/preparation_base.R"))
```

```{r, message = F, warning = F, eval = F}
load(url("http://author/code/spatial_coda/R/data_cantons.RData"))
source(url("http://author/code/spatial_coda/R/preparation_base.R"))
```

Preparation of the data :

```{r}
Ye <- as(y_ilr, "matrix")
Xe <- as(cbind(1, x2_df[, c("diplome3_ilr1", "diplome3_ilr2",
                            "employ_ilr1", "employ_ilr2", "employ_ilr3", 
                            "employ_ilr4",
                            "age3_ilr1", "age3_ilr2", 
                            "unemp_rate", "income_rate", "voters")]),
         "matrix")
ne <- nrow(Xe)
k <- ncol(Xe)
```

Spatial weight matrix : 

```{r, warning = F, message = F}
coords_fr <- st_coordinates(st_centroid(contours_occitanie_no0))
W_listw <- nb2listw(knn2nb(knearneigh(coords_fr[, 1:2], 5)), 
                    style = "W")
W_dep <- listw2mat(W_listw)
```

We represent the vote shares:

```{r, fig.width = 15, fig.height = 5, warning = F, message = F}
R <- c(0, 0)
G <- c(1, 0)
B <- c(0.5, sqrt(3) / 2)
Y <- as(ilrInv(Ye), "matrix")
# create a grid of shares for Y
my_don_x_new <- Y[, 1] * R[1] + Y[, 2] * G[1] + Y[, 3] * B[1] 
my_don_y_new  <- Y[, 1] * R[2] + Y[, 2] * G[2] + Y[, 3] * B[2] 

# create a grid of shares for Y
#pdf(file = "figures/vote_shares_cantons.pdf", width = 11, height = 4)
op <- par(oma = c(1, 1, 0, 1), mar = c(0.7, 0.7, 0.1, 0.7),
          mfrow = c(1, 2))
plot(rbind(R, G, B), 
     xaxt = "n", 
     yaxt = "n", 
     frame = F, 
     pch = 16, 
     type = "n",
     xlab = "", 
     ylab = "", 
     asp = 1, 
     ylim = c(-0.1, sqrt(3)/2 +0.05)
)


lines(c(0, 1), c(0, 0), col = "black")
lines(c(0, 0.5), c(0, sqrt(3)/2), col = "black")
lines(c(1, 0.5), c(0, sqrt(3)/2), col = "black")

# XR lines
for (k in 0:5) {
  lines(c(k/10, 1-k/10), c(k/5 * B[2], k/5 * B[2]), lty = 2, lwd = 0.8)
  text(1-k/10, k/5 * B[2], 2 * k/10, pos = 4, cex = 0.7)
}

# Left lines
for (k in 0:5) {
  lines(c(k/5, 0.5 + k * 1/10), c(0, B[2] - k/5 * B[2]), lty = 2, lwd = 0.8)
  text(k/5, 0, 2 * k/10, pos = 1, cex = 0.7, srt = 70)
}

# Right lines 
for (k in 0:5) {
  lines(c(k * 1 / 10, k/5), c(k/5 * B[2], 0), lty = 2, lwd = 0.8)
  text(k * 1 / 10, k/5 * B[2], 1 - 2 * k/10, pos = 2, cex = 0.7, srt = -50)
}

points(my_don_x_new, my_don_y_new, pch = 16, 
       col = my_rgb_coda(Y, max_intensity = 1.5), 
       cex = 1)
text(c(0.04, 0.95, 1/2), c(0 - 0.07, -0.07, sqrt(3)/2 - 0.005), 
     c("Left", "Right", "XR"), pos = c(2, 4, 3), 
     cex = 0.8)

points(my_don_x_new[c(72, 24, 94)], my_don_y_new[c(72, 24, 94)], 
       pch = 17, 
       col = rgb(0.3, 0.3, 0.3), 
       cex = 1.7)

plot(st_geometry(contours_occitanie_no0), 
     col = my_rgb_coda(Y, max_intensity = 1.5),
     border = my_rgb_coda(Y, max_intensity = 1.5), lwd = 0.1)
plot(st_geometry(st_centroid(contours_occitanie_no0[c(72, 24, 94), ])), 
       add = T, pch = 17, 
       col = rgb(0.3, 0.3, 0.3), cex = 1.9)
coords_pts <- st_coordinates(st_geometry(st_centroid(
  contours_occitanie_no0[c(72, 24, 94), ])))
text(coords_pts[, 1], coords_pts[, 2] + 5000, 
     contours_occitanie_no0[c(72, 24, 94), ]$NOMCT,
     pos = c(3, 3, 4), cex = 0.78, col = rgb(0.2, 0.2, 0.2))
par(op)
#dev.off()
```


## Spatial CoDa Model :

```{r, eval = T, echo = F}
source(url("http://www.thibault.laurent.free.fr/code/spatial_coda/R/estimate_spatial_multi_gen_N.R"))
```

```{r, eval = F, echo = T}
source(url("http://author/code/spatial_coda/R/estimate_spatial_multi_gen_N.R"))
```


```{r}
W_listw <- nb2listw(knn2nb(knearneigh(coords_fr[, 1:2], 5)), 
                    style = "W")
W_dep <- listw2mat(W_listw)
(res_sar_N <- estimate_spatial_multi_gen_N(Y = Ye, X = Xe, W = W_dep, 
                  method = "s2sls",
                  ind_RHO = matrix(c(T, T, T, T), 2, 2), compute_sd = T)
)
```

Semi-elasticities:

```{r}
my_beta <- res_sar_N$est_beta
elast_sp <- sp_semi_elast_Ycompo(Xe, my_beta, index_b = 11, method = "exact",
                                 W = W_dep, 
                                 RHO = res_sar_N$est_RHO)
```


We represent the semi-elasticities in the three cantons chosen in the article (Saint-Gaudens, Lot et Palanges, Agde): 

```{r}
# fix the bounds of the colors
elas_k <- c(elast_sp[c(72, 24, 94), , 1], 
            elast_sp[c(72, 24, 94), , 2], 
            elast_sp[c(72, 24, 94), , 3])

pos_class <- classInt::classIntervals(elas_k[elas_k>0], 6, "kmeans")
neg_class <- classInt::classIntervals(elas_k[elas_k<0], 6, "kmeans")
# we plot the se on the map 
rwb <- colorRampPalette(colors = c("blue", "white", "red"))
```


```{r, fig.width = 14, fig.height=10}
require(cartography)
# pdf("figures/elasticites_spatial_canton_123.pdf", width = 8.5, height = 5.5)
op <- par(mfrow = c(3, 3), oma = c(0, 1, 0, 0), mar = c(0, 0, 1, 0))
i <- 1
for(k in c(72, 24, 94)) {
  contours_occitanie_no0$elas_1 <- elast_sp[k, , 1]
  contours_occitanie_no0$elas_2 <- elast_sp[k, , 2]
  contours_occitanie_no0$elas_3 <- elast_sp[k, , 3]
  # we add one class for positive and negative
  choroLayer(contours_occitanie_no0, var = "elas_1", 
             breaks = c(neg_class$brks[1:6], pos_class$brks[2:7]),
             col = rwb(11),legend.pos = "n", legend.values.rnd = 0, 
             legend.title.txt = "se (left)", border = rgb(0.5, 0.5, 0.5), lwd = 0.2)
  if (i == 1) {
    title("se Left")
text(390000, 2000000, "Impacts do to a change \n at Saint-Gaudens", 
     pos = 2, cex = 1, srt = 90)
  }
  if (i == 2) {
    text(390000, 2000000, "Impacts do to a change \n at Lot et Palanges", 
     pos = 2, cex = 1, srt = 90)
  }
  if (i == 3) {
    text(390000, 2000000, "Impacts do to a change \n at Agde", 
     pos = 2, cex = 1, srt = 90)
  }
  choroLayer(contours_occitanie_no0, var = "elas_2", 
             breaks = c(neg_class$brks[1:6], pos_class$brks[2:7]),
             col = rwb(11),legend.pos = "n", legend.values.rnd = 0, 
             legend.title.txt = "se (right)", border = rgb(0.5, 0.5, 0.5), lwd = 0.2)
  if (i == 1)
    title("se Right")
  choroLayer(contours_occitanie_no0, var = "elas_3", 
             breaks = c(neg_class$brks[1:6], pos_class$brks[2:7]),
             col = rwb(11),legend.pos = c(760000, 1800000), 
             legend.values.rnd = 2, legend.title.txt = "se", 
             border = rgb(0.5, 0.5, 0.5), lwd = 0.2, legend.values.cex = 0.8)
  if (i == 1)
    title("se Extreme Right")
  i <- i + 1
}
par(op)
# dev.off()
```

We zoom on the three cantons :

```{r, warning = F}
coords_s <- st_coordinates(st_centroid(contours_occitanie_no0))
```


```{r, message = F, warning = F, fig.height = 5, fig.width= 5}
# pdf("figures/zoom_canton_2.pdf", width = 5, height = 5)
par(oma = c(1, 1, 0, 1), mar = c(0.7, 0.7, 0.1, 0.7))
map_elas_ternary(coords_s, add = F, row_to_plot = 72,
                  X = Xe, b_star = my_beta, index_b = 11, W = W_dep,
                 type = 1, scale_ternary = TRUE,
                  RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), delta = 0.1,
                   ternary_unit = T, range = 1, 
                 cex.pts = 0.2,  length = 0.05, 
                 cex.legend = 0.6, lwd = 1.6,
                 legend.ternary = c("Left", "Right", "XR"))
text(0.4, 0.25, "Saint-Gaudens", cex = 0.6, col = rgb(0.2, 0.2, 0.2), pos = 3)
map_elas_ternary(coords_s, add = T, row_to_plot = 24,
                  X = Xe, b_star = my_beta, index_b = 11, W = W_dep,
                 type = 1, scale_ternary = FALSE,
                  RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), delta = 0.1,
                   ternary_unit = T, range = 1, 
                 cex.pts = 0.2,  length = 0.05, 
                 cex.legend = 0.6, lwd = 1.6,
                 legend.ternary = c("", "", ""))
text(0.53, 0.42, "Agde", cex = 0.6, col = rgb(0.2, 0.2, 0.2), pos = 3)
map_elas_ternary(coords_s, add = T, row_to_plot = 94,
                  X = Xe, b_star = my_beta, index_b = 11, W = W_dep,
                 type = 1, scale_ternary = FALSE,
                  RHO = matrix(c(0.65, 0, 0.18, 0.63), 2, 2), delta = 0.1,
                   ternary_unit = T, range = 1, 
                 cex.pts = 0.2,  length = 0.05, 
                 cex.legend = 0.6, lwd = 1.6,
                 legend.ternary = c("", "", ""))
text(0.68, 0.16, "Lot et Palanges", cex = 0.6, col = rgb(0.2, 0.2, 0.2), pos = 3)
legend("topright", legend = c("direct", "indirect"), pch = c(NA, NA),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, box.lwd = -1) #normal legend. Do not plot line
par(font = 5) #change font to get arrows
legend("topright", legend = c("        ", "          "), pch = c(174, 174),
       lwd = 3, col = c("#E16A86", "#009ADE"),
       cex = 0.9, lty = NA, bty = "n", box.lwd = -1) 
# dev.off()
```


We summarize the results into direct, indirect and total effects:

```{r}
spatial_direct <- apply(elast_sp, 3, function(x) diag(x))
spatial_total <- apply(elast_sp, c(1, 3), sum)
spatial_indirect <- spatial_total - spatial_direct
colnames(spatial_direct) <- colnames(spatial_total) <- 
  colnames(spatial_indirect) <- c("Left", "Right", "XR")
```

We now plot the individual barplot 

```{r, fig.width = 5, fig.height = 5}
# pdf("figures/bar_local_cantons.pdf", width = 6, height = 3.5)
par(oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
plot(st_geometry(contours_occitanie_no0), 
     border = rgb(0.8, 0.8, 0.8), lwd = 0.1)
map_elas_bar(spatial_direct, spatial_indirect, spatial_total, coords_s, add = T,
             index_to_plot = c(72, 24, 94), plot_scale = F,       
             q4 = my_rgb_coda(as(rbind(c(1, 0, 0), c(0, 1, 0), c(0, 0, 1)), "matrix"), 
                    max_intensity = 1.5),
         width_bar = 2, max_bar = 12, cex.legend = 0.6, x.legend = "topleft",
         label_s = F, round.scale = 0)
text(coords_pts[, 1], coords_pts[, 2]+10000, 
     contours_occitanie_no0[c(72, 24, 94), ]$NOMCT,
     pos = 3, cex = 0.8, col = rgb(0.1, 0.1, 0.1))
# dev.off()
```

